---
title: "SimulationVisualisationCode"
author: "Chris Cox"
date: "2026-04-09"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Load necessary packages:
```{r}
pacman::p_load(ggplot2, 
               dplyr, 
               cowplot,
               abcrf,
               here,
               tidyr,
               latex2exp,
               diptest)
```

# Source model simulation functions

```{r}
source(here("SimulationFunctions/01_eventtriggered.R"))
source(here("SimulationFunctions/02_tempotracking.R"))
source(here("SimulationFunctions/03_statedependent.R"))
source(here("CustomFunctions/functions.R"))
```


# Run simulations

```{r}
set.seed(24)

sim1 <- simulate_event_triggered(
  duration = 300, T1 = 1.0, T2 = 1.0,
  rho = 0.65, d_call = 0.2, noise_sd = 0.2, duration_noise_sd = 0.1)

sim2 <- simulate_coupled_oscillator(
  duration = 1000, freq1_init = 0.8, freq2_init = 1.2,
  K = -0.1, epsilon = 0.05, dt = 0.01,
  phase_noise_sd = 0.08, freq_noise_sd = 0.001,
  d_call = 0.2, duration_noise_sd = 0.1, ici_noise_sd = 0.05,
  initial_phase_diff = 0.3 * pi
)

sim3 <- simulate_state_dependent(
  duration = 1000, lambda_active = 1.0, lambda_listening = 0.05,
  lambda_refractory = 0.001,
  p_active_to_listening_base = 0.002, p_listening_to_active_base = 0.02,
  p_active_to_listening_partner = 0.3, p_listening_to_active_partner = 0.005,
  refrac_dur = 0.3, partner_window = 0.5,
  d_call = 0.2, duration_noise_sd = 0.1, dt = 0.01
)
```

# Mechanistically Distinct Coordination Strategies (Figure 1)

## Row 1, Timeline Plots

```{r}
timeline1 <- make_timeline(sim1, time = " ")
timeline2 <- make_timeline(sim2, l_position = "none")
timeline3 <- make_timeline(sim3, l_position = "none", time = " ")

timeline_row <- cowplot::plot_grid(timeline1, NULL,
                                   timeline2, NULL,
                                   timeline3, NULL,
                                   ncol = 6,
                                   rel_widths = c(0.3, 0.05, 0.3, 0.05, 0.3, 0.05))
```

## Row 2, Inter-Call Interval Histograms

```{r}
ici1 <- make_ici(sim1, bins = 30, toggle_x = " ",
                 x_lower_lim = 0, x_upper = 4, breaks = 1)

ici2 <- make_ici(sim2, , bins = 20, x_upper_lim = 1.5, x_lower_lim = 0.5, breaks = 0.5)

ici3 <- make_ici(sim3, bins = 30, toggle_x = " ", breaks = 1, x_upper = 10)

ici_row <- cowplot::plot_grid(ici1, NULL,
                              ici2, NULL,
                              ici3, NULL,
                              ncol = 6,
                              rel_widths = c(0.3, 0.05, 0.3, 0.05, 0.3, 0.05))
ici_row
```

## Row 3, Phase Dynamics Plots

```{r}
diag1 <- make_phase_dynamics_m1(sim1, window)
diag2 <- make_phase_dynamics_m2(sim2, window)
diag3 <- make_phase_dynamics_m3(sim3, window)

diag_row <- cowplot::plot_grid(diag1, NULL,
                               diag2, NULL,
                               diag3, NULL,
                               ncol = 6,
                               rel_widths = c(0.3, 0.05, 0.3, 0.05, 0.3, 0.05))
```

## Assemble Figure 1

```{r}
main_rows <- cowplot::plot_grid(
  timeline_row,
  NULL,
  ici_row,
  NULL,
  diag_row,
  NULL,
  ncol = 1,
  rel_heights = c(0.3, 0.05, 0.3, 0.05, 0.3, 0.03)
)

full_figure <- cowplot::plot_grid(
  header_row,
  cowplot::plot_grid(row_labels, main_rows, ncol = 2, rel_widths = c(0.05, 0.95)),
  ncol = 1,
  rel_heights = c(0.03, 0.88)
)

full_figure

# Save outputs
#ggsave("figure_combined.png", full_figure, width = 14.1, height = 10, dpi = 300, bg = "white")
#ggsave(file = "figure_combined.pdf", plot = full_figure, width = 14, height = 15, bg = "white")
#ggsave("figure_combined.tiff", full_figure, width = 14.1, height = 10, dpi = 300, bg = "white")
```

# Cross-Taxonomic Distribution of Models (Figure 2)

```{r}
tax_full <- read.csv(here('tax_full.csv')) %>%
  mutate(model = factor(model, levels = c("State-Dependent", "Tempo-Tracking", "Event-Triggered"))) %>%
  mutate(taxon = factor(taxon, levels = c("Orthoptera", "Anurans", "Aves", "Primates", "Cetaceans", "Pinnipeds", "Herpestids")))


p_tax <- ggplot(tax_full, aes(x = taxon, y = model, fill = n)) +
  geom_tile(color = "white", linewidth = 1.5) +
  geom_text(aes(
    label = ifelse(n > 0, as.character(n), ""),
    color = ifelse(n >= 3, "white", "black")   # inside aes()
  ),
  size = 10, fontface = "bold") +
  scale_color_identity() +   # tells ggplot to use the colour values literally
  scale_fill_gradient(
    low    = "#EFF3FF",
    high   = "#08306B",
    name   = "Studies (N)",
    breaks = c(0, 3, 6, 9),
    labels = c("0", "3", "6", "9"),
    limits = c(0, 9),
    guide  = guide_colorbar(
      barwidth       = 1.5,
      barheight      = 15,
      ticks          = FALSE,
      title.position = "top",
      title.hjust    = 0.5,
      label.hjust    = 0
    )
  ) +
  scale_x_discrete(position = "bottom") +
  coord_fixed() +
  labs(
    x     = "Taxonomic Group",
    y     = "Model Type",
    #title = "Cross-Taxonomic Distribution of Formal Coordination Models"
  ) +
  theme_minimal(base_size = 13, base_family = "Helvetica") +
  theme(
    plot.title   = element_text(face = "bold", size = 20, hjust = 0.5, color = "black"),
    axis.text.x  = element_text(face = "bold", size = 15, color = "black"),
    axis.text.y  = element_text(face = "bold", size = 15, color = "black"),
    axis.title.y = element_text(face = "bold", size = 15, color = "black"),
    axis.title.x = element_text(face = "bold", size = 15, color = "black", margin = margin(t = 15)),
    panel.grid   = element_blank(),
    legend.position = "right",
    legend.title = element_text(size = 16, face = "bold"),  # Bigger title
    legend.text = element_text(size = 15),  # Bigger numbers
  )

p_tax_title <- make_panel_title("Cross-Taxonomic Distribution of Formal Coordination Models")

heatmap_studies <- plot_grid(
  p_tax_title, 
  p_tax, 
  ncol = 1,
  rel_heights = c(0.08, 0.93))

heatmap_studies

#ggsave(plot = heatmap_studies, file = here('heatmap_studies.pdf'), height = 5, width = 15.5)
#ggsave(plot = heatmap_studies, file = here('heatmap_studies.tiff'), height = 5, width = 15.5)
#ggsave(plot = heatmap_studies, file = here('heatmap_studies.png'), height = 5, width = 15.5)
```

# Model Comparison (Figure 3)

```{r}
# ABC Model Selection + Posterior Predictive

# What this code block does:
# 1. Simulate reference datasets from each model with random parameters
#    (this is the "prior predictive", exploring what each model can produce)
# 2. Compute 8 summary statistics for each simulated dataset
# 3. Train a random forest (abcrf) to classify which model generated each dataset
# 4. Apply to observed data to get posterior model probabilities
# 5. Find closest reference simulations to observed data (rejection ABC)
# 6. Resimulate from those parameters to get posterior predictive distributions

stat_names <- c("mean_ici", "sd_ici", "bimod", "prop_short", "dip", "phase_dist", "phase_var", "phase_acf1")
```

## Step 1: Compute observed statistics

```{r}
# Function takes call times from two individuals, returns 8 statistics
obs_event <- compute_8_stats(sim1$voc_times_1, sim1$voc_times_2)
obs_tempo  <- compute_8_stats(sim2$voc_times_1, sim2$voc_times_2)
obs_state  <- compute_8_stats(sim3$voc_times_1, sim3$voc_times_2)

df_obs <- rbind(event = obs_event, tempo = obs_tempo, state = obs_state)

colnames(df_obs) <- stat_names

print(round(df_obs, 3))
```

## Step 2: Build reference table

```{r}
# For each model: sample random parameters from prior, simulate,
# compute statistics, store parameters + statistics

n_sim <- 300   # simulations per model
dur   <- 1000    # simulation duration in seconds

model_labels <- character(0) # storage for abcrf, model labels
stats_matrix <- matrix(NA, nrow = 0, ncol = 8) # storage for abcrf, statistics
```

### Event-Triggered

```{r}
csv_event <- "ref_tables/ref_event.csv"

# --- Event-Triggered ---
for (i in 1:n_sim) {
  T        <- runif(1, 0.5, 2.0)
  rho      <- runif(1, 0.3, 0.9)
  d_call   <- runif(1, 0.05, 0.5)
  noise_sd <- runif(1, 0.01, 0.15)

  sim <- simulate_event_triggered(
    duration = dur, T1 = T, T2 = T,
    rho = rho, d_call = d_call,
    noise_sd = noise_sd, duration_noise_sd = noise_sd
  )
  stats <- compute_8_stats(sim$voc_times_1, sim$voc_times_2)

  if (!any(is.na(stats))) {
    row <- data.frame(
      T=T, rho=rho, d_call=d_call, noise_sd=noise_sd,
      s1=stats[1], s2=stats[2], s3=stats[3], s4=stats[4],
      s5=stats[5], s6=stats[6], s7=stats[7], s8=stats[8]
    )
    write_or_append(row, csv_event)
    model_labels <- c(model_labels, "Event-Triggered")
    stats_matrix <- rbind(stats_matrix, stats)
  }
}

ref_event <- read.csv(csv_event)
```

### Tempo-Tracking

```{r}
csv_tempo <- "ref_tables/ref_tempo.csv"

# --- Tempo-Tracking ---
for (i in 1:n_sim) {
  K              <- runif(1, 0.01, 0.2)
  epsilon        <- runif(1, 0.01, 0.2)
  freq1          <- runif(1, 0.5, 2.0)
  freq2          <- runif(1, 0.5, 2.0)
  phase_noise_sd <- runif(1, 0.01, 0.2)

  sim <- simulate_coupled_oscillator(
    duration = dur, freq1_init = freq1, freq2_init = freq2,
    K = -K, epsilon = epsilon, dt = 0.02,
    phase_noise_sd = phase_noise_sd, freq_noise_sd = 0.001,
    d_call = 0.2, duration_noise_sd = 0.1, ici_noise_sd = 0.05,
    initial_phase_diff = runif(1, 0, pi)
  )
  stats <- compute_8_stats(sim$voc_times_1, sim$voc_times_2)

  if (!any(is.na(stats))) {
    row <- data.frame(
      K=K, epsilon=epsilon, freq1=freq1, freq2=freq2,
      phase_noise_sd=phase_noise_sd,
      s1=stats[1], s2=stats[2], s3=stats[3], s4=stats[4],
      s5=stats[5], s6=stats[6], s7=stats[7], s8=stats[8]
    )
    write_or_append(row, csv_tempo)
    model_labels <- c(model_labels, "Tempo-Tracking")
    stats_matrix <- rbind(stats_matrix, stats)
  }
}

ref_tempo <- read.csv(csv_tempo)
```

### State-Dependent

```{r}
csv_state <- "ref_tables/ref_state.csv"

# --- State-Dependent ---
for (i in 1:n_sim) {
  lambda_active <- runif(1, 0.5, 3.0)
  p_on          <- runif(1, 0.005, 0.1)
  p_off         <- runif(1, 0.001, 0.02)

  sim <- simulate_state_dependent(
    duration = dur,
    lambda_active = lambda_active,
    lambda_listening = 0.05, lambda_refractory = 0.001,
    p_active_to_listening_base = p_off,
    p_listening_to_active_base = p_on,
    p_active_to_listening_partner = 0.3,
    p_listening_to_active_partner = 0.005,
    refrac_dur = 0.3, partner_window = 0.5,
    d_call = 0.2, duration_noise_sd = 0.1, dt = 0.01
  )
  stats <- compute_8_stats(sim$voc_times_1, sim$voc_times_2)

  if (!any(is.na(stats))) {
    row <- data.frame(
      lambda_active=lambda_active, p_on=p_on, p_off=p_off,
      s1=stats[1], s2=stats[2], s3=stats[3], s4=stats[4],
      s5=stats[5], s6=stats[6], s7=stats[7], s8=stats[8]
    )
    write_or_append(row, csv_state)
    model_labels <- c(model_labels, "State-Dependent")
    stats_matrix <- rbind(stats_matrix, stats)
  }
}

ref_state <- read.csv(csv_state)
```

## Step 3: Train random forest (abcrf)

```{r}
colnames(stats_matrix) <- stat_names

ref_table <- data.frame(model = as.factor(model_labels), stats_matrix)

model_rf <- abcrf(model ~ ., data = ref_table, ntree = 500, paral = TRUE) # predict model identity from all 8 summary statistics

print(model_rf$model.rf$confusion.matrix)
cat("Out-of-bag error rate:", round(model_rf$prior.err, 3), "\n")
```

## Step 4: Model selection based on "observed" data

```{r}
# The near-perfect classification here is validation only (since datasets are simulated); 
# the real rest will be applying this to observed data where the generative mechanism is unknown.
obs_event_df <- as.data.frame(t(obs_event)); colnames(obs_event_df) <- stat_names
obs_tempo_df <- as.data.frame(t(obs_tempo)); colnames(obs_tempo_df) <- stat_names
obs_state_df <- as.data.frame(t(obs_state)); colnames(obs_state_df) <- stat_names

pred_event <- predict(model_rf, obs=obs_event_df, training=ref_table)
pred_tempo <- predict(model_rf, obs=obs_tempo_df, training=ref_table)
pred_state <- predict(model_rf, obs=obs_state_df, training=ref_table)

# Extract variable importance, dropping LD1/LD2:
imp <- model_rf$model.rf$variable.importance %>%
  tibble::enframe() %>%
  filter(!name %in% c("LD1", "LD2")) %>%
  tibble::deframe()

print(round(sort(imp, decreasing=TRUE), 3))
```

## Step 5: Find closest reference simulations = approximate posterior

```{r}
# Rejection ABC: keep the n_best simulations whose summary statistics are closest
# to the observed data (Euclidean Distance) — their parameters form the approximate posterior.

stat_cols <- paste0("s", 1:8)
n_best    <- 50

# Event model on each dataset
best_event_on_event <- ref_event[order(compute_dist(ref_event[,stat_cols], obs_event))[1:n_best], ]
best_event_on_tempo <- ref_event[order(compute_dist(ref_event[,stat_cols], obs_tempo))[1:n_best], ]
best_event_on_state <- ref_event[order(compute_dist(ref_event[,stat_cols], obs_state))[1:n_best], ]

# Tempo model on each dataset
best_tempo_on_event <- ref_tempo[order(compute_dist(ref_tempo[,stat_cols], obs_event))[1:n_best], ]
best_tempo_on_tempo <- ref_tempo[order(compute_dist(ref_tempo[,stat_cols], obs_tempo))[1:n_best], ]
best_tempo_on_state <- ref_tempo[order(compute_dist(ref_tempo[,stat_cols], obs_state))[1:n_best], ]

# State model on each dataset
best_state_on_event <- ref_state[order(compute_dist(ref_state[,stat_cols], obs_event))[1:n_best], ]
best_state_on_tempo <- ref_state[order(compute_dist(ref_state[,stat_cols], obs_tempo))[1:n_best], ]
best_state_on_state <- ref_state[order(compute_dist(ref_state[,stat_cols], obs_state))[1:n_best], ]
```

## Step 6: Posterior predictive ICI simulations

```{r}
### Resimulate from the closest parameter sets
n_ppc <- 10

# Row 1: Event data — all three models
ppc_event_event <- simulate_ici_event(best_event_on_event, x_upper=5)
ppc_tempo_event <- simulate_ici_tempo(best_tempo_on_event, x_upper=5)
ppc_state_event <- simulate_ici_state(best_state_on_event, x_upper=5)

# Row 2: Tempo data — all three models
ppc_event_tempo <- simulate_ici_event(best_event_on_tempo, x_upper=2)
ppc_tempo_tempo <- simulate_ici_tempo(best_tempo_on_tempo, x_upper=2)
ppc_state_tempo <- simulate_ici_state(best_state_on_tempo, x_upper=2)

# Row 3: State data — all three models
ppc_event_state <- simulate_ici_event(best_event_on_state, x_upper=8)
ppc_tempo_state <- simulate_ici_tempo(best_tempo_on_state, x_upper=8)
ppc_state_state <- simulate_ici_state(best_state_on_state, x_upper=8)
```

## Step 7: Plot posterior predictive band + observed density

```{r}
# Diagonal panels (correct model) shown with full color
# Off-diagonal panels (wrong model) shown with muted color

# Row 1: Event data observed
n_ppc <- 10

# Row 1: Event data — all three models
p11 <- make_ppc_panel(ppc_event_event, sim1$voc_times_1, sim1$voc_times_2,
                      col_event, "Event model | Event data", 3, TRUE,
                      show_y = F)

p12 <- make_ppc_panel(ppc_tempo_event, sim1$voc_times_1, sim1$voc_times_2,
                      col_tempo, "Tempo model | Event data", 5, FALSE)
p13 <- make_ppc_panel(ppc_state_event, sim1$voc_times_1, sim1$voc_times_2,
                      col_state, "State model | Event data", 5, FALSE)

# Row 2: Tempo data observed
p21 <- make_ppc_panel(ppc_event_tempo, sim2$voc_times_1, sim2$voc_times_2,
                      col_event, "Event model | Tempo data", 2, FALSE)

p22 <- make_ppc_panel(ppc_tempo_tempo, sim2$voc_times_1, sim2$voc_times_2,
                      col_tempo, "Tempo model | Tempo data", 5, TRUE,
                      show_y = F)

p23 <- make_ppc_panel(ppc_state_tempo, sim2$voc_times_1, sim2$voc_times_2,
                      col_state, "State model | Tempo data", 2, FALSE)

# Row 3: State data observed
p31 <- make_ppc_panel(ppc_event_state, sim3$voc_times_1, sim3$voc_times_2,
                      col_event, "Event model | State data", 8, FALSE,
                      show_y = T)
p32 <- make_ppc_panel(ppc_tempo_state, sim3$voc_times_1, sim3$voc_times_2,
                      col_tempo, "Tempo model | State data", 8, FALSE,
                      show_y = T)
p33 <- make_ppc_panel(ppc_state_state, sim3$voc_times_1, sim3$voc_times_2,
                      col_state, "State model | State data", 8, TRUE,
                      show_y = T)
```

## Step 8: Assemble 3x3 grid with row and column labels

```{r}
col_labels <- c("Event-Triggered",
                "Tempo-Tracking",
                "State-Dependent")

# Add column headers
header <- plot_grid(
  ggplot() + annotate("text", x=0.5, y=0.5, label=col_labels[1],
                      fontface="bold", size=13, color=col_event) +
    theme_void(),
  ggplot() + annotate("text", x=0.5, y=0.5, label=col_labels[2],
                      fontface="bold", size=13, color=col_tempo) +
    theme_void(),
  ggplot() + annotate("text", x=0.5, y=0.5, label=col_labels[3],
                      fontface="bold", size=13, color=col_state) +
    theme_void(),
  ncol=3
)

row1 <- plot_grid(p11, NULL, p12, NULL, p13, NULL, ncol=6, rel_widths = c(0.5, 0.05, 0.5, 0.05, 0.5, 0.05))
row2 <- plot_grid(p21, NULL, p22, NULL, p23, NULL, ncol=6, rel_widths = c(0.5, 0.05, 0.5, 0.05, 0.5, 0.05))
row3 <- plot_grid(p31, NULL, p32, NULL, p33, NULL, ncol=6, rel_widths = c(0.5, 0.05, 0.5, 0.05, 0.5, 0.05))

full_fig <- plot_grid(row1, NULL, row2, NULL, row3, ncol=1, rel_heights=c(1, 0.1, 1, 0.1, 1))
full_fig
```

## Mechanistic Assumption and Diagnostic Predictions (Figure 3)

### Event-Triggered (Inhibitory Reset)

```{r}
# Three model "cards" side by side with full equations,
# mechanistic description, and diagnostic predictions

p_event <- make_model_card(
  color    = col_event,
  title    = "Event-Triggered",
  subtitle = "Discrete stimulus-triggered corrections\nto an endogenous oscillator",
  equations = list(
    TeX("$\\phi_i(t^+) = \\phi_i(t^-) + \\rho \\cdot (\\pi - \\phi_i(t^-))$"),
    TeX("when partner $j$ calls at time $t$"),
    TeX("$ICI \\sim N(T,\\, \\sigma^2)$ or $N(2(d + \\rho T),\\, \\sigma^2)$"),
    TeX("$\\dot{\\phi}_i = 2\\pi / T_i + \\epsilon(\\bar{T} - T_i)$")
  ),
  eq_y = c(0.7, 0.62, 0.53, 0.44),
  predictions = list(
    "Inter-Call Intervals"    = "Bimodal",
    "Phase Persistence"   = "Negative",
    "Phase Variance"      = "High (jitter)",
    "Convergence"         = "Discrete resets"
  )
)
```

### Tempo-Tracking (Kuramoto Coupled Oscillator)

```{r}
p_tempo <- make_model_card(
  color    = col_tempo,
  title    = "Tempo-Tracking",
  subtitle = "Continuous mutual adjustment\nof calling frequencies",
  equations = list(
    TeX("$\\dot{\\phi}_i = \\omega_i + K \\cdot \\sin(\\phi_j - \\phi_i)$"),
    TeX("$\\dot{\\omega}_i = \\epsilon \\cdot (\\bar{\\omega} - \\omega_i)$"),
    TeX("$\\phi^* = \\pi$ (antiphase equilibrium, $K < 0$)"),
    TeX("$ICI \\sim N(1/\\omega^*,\\, \\sigma^2_{\\phi})$")
  ),
  eq_y = c(0.7, 0.61, 0.52, 0.43),
  predictions = list(
    "Inter-Call Intervals"    = "Unimodal, tight",
    "Phase Persistence"   = "Positive",
    "Phase variance"      = "Low (converged)",
    "Convergence"         = "Smooth, gradual"
  )
)
```

### State-Dependent

```{r}
p_state <- make_model_card(
  color    = col_state,
  title    = "State-Dependent",
  subtitle = "Probabilistic behavioral gating\nacross internal states",
  equations = list(
    TeX("$S_i(t) \\in \\{active,\\; listening,\\; refractory\\}$"),
    TeX("$P(call \\,|\\, S_i = active) = \\lambda_{active} \\cdot dt$"),
    TeX("$P(S_i: active \\to listening \\,|\\, j\\; calls) = p_{off}$"),
    TeX("$ICI \\sim Exp(\\lambda_{active})$ within bout")
  ),
  eq_y = c(0.7, 0.61, 0.52, 0.43),
  predictions = list(
    "Inter-Call Intervals"    = "Exponential",
    "Phase Persistence"   = "Near zero",
    "Phase variance"      = "High (scatter)",
    "Convergence"         = "None (stochastic)"
  )
)
```

### Assemble cards into panel

```{r}
p_cards <- plot_grid(p_event, p_tempo, p_state,
                   ncol=3, rel_widths=c(1, 1, 1))
```

## From Statistical Fingerprints to Model Inference (Figure 3)

### Row 1: Condensed fingerprint panels 

```{r}
# Rebuild ICI panels at smaller size
ici1_mini <- make_ici(sim1, bins = 30, toggle_x = " ",
                      x_lower_lim = 0, x_upper = 4, breaks = 1) +
  theme(axis.title.y = element_text(size = 12, color = "black", angle = 90)) +
  labs(y = "Density")

ici2_mini <- make_ici(sim2, bins = 20, x_upper_lim = 1.5,
                      x_lower_lim = 0.5, breaks = 0.5) +
  labs(title = " ")

ici3_mini <- make_ici(sim3, bins = 30, toggle_x = " ",
                      breaks = 1, x_upper = 10) +
  labs(title = " ")

diag1_mini <- make_phase_dynamics_m1(sim1, window) +
  labs(title = " ") +
  theme(axis.title.y = element_text(size = 12, color = "black", angle = 90),
        panel.grid.major.y = element_blank()) +
  labs(y = "Phase")

diag2_mini <- make_phase_dynamics_m2(sim2, window) +
  labs(title = " ") + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank()) +
  labs(x = "Time (s)")

diag3_mini <- make_phase_dynamics_m3(sim3, window) +
  labs(title = " ") +
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank())

# Stack ICI + phase per model vertically
col_event_panel <- plot_grid(ici1_mini, diag1_mini,
                             ncol = 1, rel_heights = c(1, 1))
col_tempo_panel <- plot_grid(ici2_mini, diag2_mini,
                             ncol = 1, rel_heights = c(1, 1))
col_state_panel <- plot_grid(ici3_mini, diag3_mini,
                             ncol = 1, rel_heights = c(1, 1))

# Make coloured headers:
headers <- plot_grid(
  make_colored_col_header("Event-Triggered",  col_event),
  make_colored_col_header("Tempo-Tracking",   col_tempo),
  make_colored_col_header("State-Dependent",  col_state),
  ncol = 3
)

# Stack ICI + phase per model horizontally
row1_panels <- plot_grid(col_event_panel, 
                         col_tempo_panel, 
                         col_state_panel,
                         ncol = 3, nrow = 1)

diagnostic_sigs <- make_panel_title("i) Diagnostic Signatures", size = 4)

row1 <- plot_grid(
  diagnostic_sigs,
  headers,
  row1_panels,
  ncol = 1,
  rel_heights = c(0.08, 0.07, 0.93)
)
```


### Row 2, scatter plots, lower-left: 

```{r}
set.seed(42)

ref_sample <- ref_table[
  c(sample(which(ref_table$model == "Event-Triggered"), 300),
    sample(which(ref_table$model == "Tempo-Tracking"),  300),
    sample(which(ref_table$model == "State-Dependent"), 300)), ]

# Observed points for all 8 statistics
obs_df <- data.frame(
  model      = c("Event-Triggered", "Tempo-Tracking", "State-Dependent"),
  mean_ici   = c(obs_event[1], obs_tempo[1], obs_state[1]),
  sd_ici     = c(obs_event[2], obs_tempo[2], obs_state[2]),
  bimod      = c(obs_event[3], obs_tempo[3], obs_state[3]),
  prop_short = c(obs_event[4], obs_tempo[4], obs_state[4]),
  dip        = c(obs_event[5], obs_tempo[5], obs_state[5]),
  phase_dist = c(obs_event[6], obs_tempo[6], obs_state[6]),
  phase_var  = c(obs_event[7], obs_tempo[7], obs_state[7]),
  phase_acf1 = c(obs_event[8], obs_tempo[8], obs_state[8])
)

d1 <- make_scatter("prop_short", "phase_acf1",
                   "Prop. Short Intervals", "Phase Persistence")

d2 <- make_scatter("sd_ici", "phase_dist",
                   "Interval Variability", "Phase Distance")

d3 <- make_scatter("phase_var", "bimod",
                   "Phase variance", "Skewness-Kurtosis")

d4 <- make_scatter("dip", "mean_ici",
                   "Dip statistic", "Mean Interval",
                   show_legend = TRUE)

# Extract shared legend from d4
shared_legend <- get_legend(d4 +
                              theme(legend.position = "bottom",
                                    legend.text = element_text(size = 15),
                                    legend.key.size = unit(1.2, "lines")))

scatters_2x2a <- plot_grid(
  d1 + theme(legend.position = "none"),
  NULL,
  d2 + theme(legend.position = "none"),
  ncol = 3,
  rel_widths = c(1, 0.05, 1)
)

scatters_2x2b <- plot_grid(
  d3 + theme(legend.position = "none"),
  NULL,
  d4 + theme(legend.position = "none"),
  ncol = 3,
  rel_widths = c(1, 0.05, 1)
)

scatters_2x2 <- plot_grid(scatters_2x2a, 
                          NULL,
                          scatters_2x2b,
                          nrow = 3,
                          rel_heights = c(1, 0.2, 1))


panel_D_title <- make_panel_title("ii) Parameters Cluster by Mechanism", size = 4)

panel_D <- plot_grid(
  panel_D_title,
  scatters_2x2,
  shared_legend,
  ncol = 1,
  rel_heights = c(0.05, 0.93, 0.08)
)

panel_D
```

### Row 2, variable importance, lower-right: 

```{r}
imp_df <- data.frame(
  statistic  = names(imp),
  importance = as.numeric(imp)
)

imp_df <- imp_df[order(imp_df$importance, decreasing = FALSE), ]

imp_df$label <- dplyr::recode(imp_df$statistic,
                              "mean_ici"   = "Mean Interval",
                              "sd_ici"     = "Interval Variability",
                              "bimod"      = "Skewness-Kurtosis",
                              "prop_short" = "Prop. Short Intervals",
                              "dip"        = "Dip Statistic",
                              "phase_dist" = "Phase Distance",
                              "phase_var"  = "Phase Variance",
                              "phase_acf1" = "Phase Persistence")

imp_df$label <- factor(imp_df$label, levels = imp_df$label)

imp_df$stat_type <- dplyr::recode(imp_df$statistic,
                                  "mean_ici"   = "Inter-Call Interval",
                                  "sd_ici"     = "Inter-Call Interval",
                                  "bimod"      = "Inter-Call Interval",
                                  "prop_short" = "Inter-Call Interval",
                                  "dip"        = "Inter-Call Interval",
                                  "phase_dist" = "Phase",
                                  "phase_var"  = "Phase",
                                  "phase_acf1" = "Phase"
)

imp_df$stat_type <- factor(imp_df$stat_type, levels = c("Inter-Call Interval", "Phase"))

p_importance <- ggplot(imp_df,
                       aes(x = label, y = importance, color = stat_type)) +
  geom_segment(aes(x = label, xend = label, y = 0, yend = importance),
               linewidth = 0.9, alpha = 0.9) +
  geom_point(size = 5) +
  scale_color_manual(
    values = c("Inter-Call Interval" = "#F9A825",
               "Phase" = "#546E7A"),
    name   = "Dimension Type"
    
  ) +
  coord_flip() +
  labs(x = NULL, y = "Variable Importance") +
  theme_classic(base_size = 10) +
  theme(
    plot.title   = element_text(face = "bold", size = 20, hjust = 0.5, color = "black"),
    axis.text.x  = element_text(size = 15, color = "black"),
    axis.text.y  = element_text(size = 15, color = "black"),
    axis.title.y = element_text(size = 15, color = "black"),
    axis.title.x = element_text(size = 15, color = "black", margin = margin(t = 15)),
    panel.grid   = element_blank(),
    legend.position = c(0.9, 0.2),
    legend.title = element_text(size = 15, face = "bold"),  # Bigger title
    legend.text = element_text(size = 15),  # Bigger numbers
  )

panel_Diii_title <- make_panel_title("iii) Parameters Driving Inference", size = 4)

p_importance_final <- plot_grid(
  panel_Diii_title,
  p_importance,
  ncol = 1,
  rel_heights = c(0.05, 0.93)
)

p_importance_final

```

### Assemble Row 2

```{r}
row2 <- plot_grid(panel_D, NULL, p_importance_final,
                  ncol = 3, rel_widths = c(1.5, 0.1, 1.0))
```

## Full Panel B:

```{r}
fig2 <- plot_grid(
  row1,
  NULL,
  row2,
  ncol = 1,
  rel_heights = c(0.5, 0.05, 1)
)

print(fig2)
```

# Assemble Full Figure 3

```{r}
main_title_cards <- make_panel_title("A) Mechanistic Assumptions and Diagnostic Predictions")
cards_row <- plot_grid(p_event, p_tempo, p_state, ncol=3, rel_widths=c(1,1,1))

row1_w_titles <- plot_grid(
  main_title_cards, 
  cards_row, 
  ncol = 1,
  rel_heights = c(0.08, 0.93))

p_B_ABC <- make_panel_title("B) From Statistical Fingerprints to Model Inference")
p_C_Model <- make_panel_title("C) Cross-Model Predictive Fit Under ABC")
row2_titles <- plot_grid(p_B_ABC, NULL, p_C_Model, nrow = 1, rel_widths = c(0.5, 0.05, 0.5))

ABC_part <- cowplot::plot_grid(fig2, NULL, full_fig, 
                               rel_widths = c(0.5, 0.1, 0.5),
                               nrow = 1)

row2_w_titles <- plot_grid(
          row2_titles, 
          ABC_part, 
          ncol = 1,
          rel_heights = c(0.05, 0.93))

final_fig <- plot_grid(row1_w_titles, NULL, row2_w_titles,
          ncol = 1, rel_heights = c(0.4, 0.05, 0.9))

#ggsave(plot = final_fig, file = 'final_fig.pdf', height = 19.5, width = 26)
#ggsave(plot = final_fig, file = 'final_fig.png', height = 19.5, width = 26)
#ggsave(plot = final_fig, file = 'final_fig.tiff', height = 19.5, width = 26)
```
